Self Referencing Sequences

### 2014-10-31 draft.tex

\subsection{Problem Setup}
Consider the sequence
\begin{align*}
1,2,2,1,1,2,1,2,2,1,2,2,1,1,2,1,1,2,2,\dots
\end{align*}
We can break up the sequence into contiguous blocks, where each block is a stretch of repeated numbers:
\begin{align*}
[1],[2,2],[1,1],[2],[1],[2,2],[1],[2,2],[1,1],[2],[1,1],[2,2],\dots
\end{align*}
The block lengths, read from left to right, reproduce the original sequence. Thus, this sequence is \emph{self-referential}. For the remainder of this paper we study properties of such sequences.

We say that this sequence is \emph{generated} by the set of numbers used in the sequence, so the above sequence is generated by $\{1,2\}$. We will refer to this set as the \emph{generating set}. Observe that it is unambiguous which number to use next. After a block of 1's, the next block must be a block of 2's. There does exist ambiguity in a sequence with generating set $\{1,2,3\}$, since after a block of 1's, ther can be a block of 2's or a block of 3's. We assign an order to the generating set and resolve the ambiguity. The order we will use will be the order of the elements presented, for example, $\{1, 2, 3 \}$ will have a order $1 \rightarrow 2 \rightarrow 3 \rightarrow 1$. Note that because order matters, the \emph{generating set} is technically not a set, but we do not call it a ``sequence" to avoid possible confusion with its corresponding self-referencing sequence.

With this rule, a self-referencing sequence generated by $\{1,2,3\}$ has blocks of 1's followed by blocks of 2's followed by blocks of 3's, followed by blocks of 1's, and so on. Thus, $\{1,2,3\}$ can generate
\begin{align*}
1,2,2,3,3,1,1,1,2,2,2,3,2,1,2,2,3,3,1,1,\dots
\end{align*}
but not
\begin{align*}
1,3,3,3,2,2,2,1,1,1,3,3,2,2,1,1,3,2,1,3,\dots
\end{align*}
as the second sequence has the numbers in the wrong order.

@@

In this paper, we address a variety of questions regarding these sequences. In Section~\ref{determinism} we show that if we are given a generating set and a starting number for the sequence, the sequence is uniquely determined. After that, we consider how the starting number for the sequence affects the long term behavior of the sequence. In particular, in Section~\ref{prefixes}, we show that there exist limits on how much two sequences with the same generating set can resemble each other if they start with different numbers. In the following Section~\ref{equivalence}, we consider the limiting behavior of such sequences, and we conjecture that the limiting behavior of a sequence is determined only by the generating set, and not the starting number.

In Section~\ref{iterativeprocess}, we take a different approach to analyzing these sequences by showing that there is an alternate way of generating them with an iterative process. In Section~\ref{substitutionrules}, we show how a large number of these sequences can actually be generated with a simple set of substitution rules. We then introduce the density problem in Section~\ref{density}, and use ideas developed in Section~\ref{iterativeprocess} and Section~\ref{substitutionrules} to answer the question for certain types of generating sets. Unfortunately, we find that the open problem of determining the density for the sequence of 1's and 2's is hard to solve using our methods.

@@

\noindent For the final draft,\\
we will improve the haikus\\
(and the whole paper).
\end{document}

>>>>

\subsection{Problem Setup}
Consider the sequence
\begin{equation}
\label{eq:onetwo}
1,2,2,1,1,2,1,2,2,1,2,2,1,1,2,1,1,2,2,\dots
\end{equation}
We can break up the sequence into contiguous blocks, where each block is a stretch of repeated numbers.
\begin{figure}
\label{fig:sequenceblocks}
\center
\includegraphics[width=0.5\linewidth]{sequenceBlocks.pdf}
\caption{Sequence broken up into contiguous blocks with lengths generating same sequence.}
\end{figure}
The block lengths, read from left to right, reproduce the original sequence (see Figure~\ref{fig:sequenceblocks}). 
Thus, this sequence is \emph{self-referential}.

The sequence~\ref{eq:onetwo} contains the numbers $1$ and $2$. In general, we want to make self-refering sequences containing other numbers as well. Once there is more than two numbers in the sequence, generating the sequence becomes a bit ambiguous. Take the simple example of a self-referencing sequence with numbers 1, 2, and 3.

If the first number is 1, then we can make the second number either a 2, or a 3. In particular, both:
\begin{equation}
1, 3, 3, 3, 2, 2, 2, \dots
\end{equation}
\begin{equation}
1, 2, 2, 1, 1, 3, 1, \dots\\
\end{equation}
are in some sense ``self-referencing'' sequences. Each time we need to specify a new block, we have a choice of two elements. We would like to avoid having a large number of choices in the generation of the sequence, and so we require that the sequence contain elements from a \emph{cycle}, and each time we start another block, the block will contain the next number in the cycle. In the case of sequence~\ref{eq:onetwo}, the cycle is $(1,2)$.

The self-referencing sequence with cycle $(1, 2, 3)$ that begins with $1$ is
\begin{figure}
\label{fig:onetwothree}
\center
\includegraphics[width=0.5\linewidth]{sequenceBlocks123.pdf}
\caption{Sequence generated by $(1,2,3)$ broken up into contiguous blocks.}
\end{figure}

Formally, we say that a sequence is \emph{generated} by the set of numbers used in the sequence. So sequence~\ref{eq:onetwo} is generated by the set $\{1, 2\}$. We will refer to the cycle used in the sequence as the \emph{generating cycle}.

Using the cycle, it is no longer ambiguous which number to use in the next block.

@@

We address a variety of questions regarding these sequences. Our main results are:
\begin{itemize}
\item The self-referencing sequence of a  given generating cycle is uniquely determined by its first number.
\item Some characterizations of sequences with the same generating cycle.
\item There exists an equivalent formulation of self-referencing sequences using substitution rules for some criteria of the generating cycles.
\item One can compute the limiting density for the numbers in the sequence if a limiting density exists, for generating cycles that have substitution rules. 
\item The limiting densities of the self-referencing sequences of generating cycles $(1,3)$ if it exists.
\end{itemize}

In Section~\ref{determinism} we show that if we are given a generating set and a starting number for the sequence, the sequence is uniquely determined. We consider how the starting number for the sequence affects the long term behavior of the sequence. In particular, in Section~\ref{prefixes}, we show that there exist limits on how much two sequences with the same generating set can resemble each other if they start with different numbers. In Section~\ref{equivalence}, we consider the limiting behavior of such sequences, and we conjecture that the limiting behavior of a sequence is determined only by the generating set, and not the starting number.

In Section~\ref{iterativeprocess}, we take a different approach to analyzing these sequences by showing that there is an alternate way of generating them with an iterative process. In Section~\ref{substitutionrules}, we show how a large number of these sequences can actually be generated with a simple set of substitution rules. We introduce the density problem in Section~\ref{density}, and use ideas developed in Section~\ref{iterativeprocess} and Section~\ref{substitutionrules} to answer the question for certain types of generating sets. Unfortunately, we find that the open problem of determining the density for the sequence of 1's and 2's is hard to solve using our methods.

@@

\noindent For the final draft,\\
we will improve the haikus\\
(and the whole paper).
\end{document}  
>>>>>>> 94de583213b09230c51ca57ed5c9897bcabc4666
